Cycles of the logistic map

### 2013-05-15 doc/pc-ijbc.tex
pc-ijbc: add logtime.gp, editing

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx_{#1}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\lam}{(\lambda)}
\newcommand{\Rlam}{(R, \lambda)}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = f(x_k) = x_{k+1}$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  the starting point of the next iteration.
%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} or a constant (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle} or a self-repeating loop,
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz}.

@@

The onset polynomial
  $P_n(R, \lambda = 1)$ for the $n$-cycles
  can be further factorized.
%
This is caused by the intersection of an $n$-cycle
  and a shorter $d$-cycle ($d|n$, $d < n$),
  which forces the two to share orbit.
%
One can show that this intersection happens only
  at the onset of the $n$-cycle, where $\lambda = 1$,
%
  and the $d$-cycle branches out, or ``bifurcates,'' $n/d$-fold
  to become the $n$-cycle.
%
The simplest example is the first bifurcation point
  at $R = 3/4$ for $d = 1, n = 2$,
  where the fixed point \refeq{xr1}
  undergoes period-doubling and becomes the 2-cycle \refeq{xr2}.
As a result, the polynomial $P_n(R, \lambda)$ from the $n$-cycle
  has to accommodate $P_d(R, \lambda')$ from the $d$-cycle as a factor,
  with $\lambda'$ being a primitive $(n/d)\,$th root of $\lambda = 1$.

If an $n$-cycle is not born out of the above branching,
  we call it \emph{original}.
The two types of cycles exist
  as separate factors of $P_n(R, \lambda = 1)$.
One can then show that the factor $S_n(R)$ responsible for the original cycles
  %or the \emph{original factor} below,
  is given by
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      %\left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      %\right]
    },
  \label{eq:origfac}
  \end{equation}
where the inner product on the denominator is carried over
   $k$ from 1 to $c$ that are coprime to $c$.

Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$,
  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
  the onset of the original cycle.
Note $R_a' = 5/4$ is excluded from $S_4(R)$
  as it is due to period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).

@@

%\begin{theorem}
The degrees in $R$ of $A_n\Rlam$, $P_n\Rlam$ and $S_n(R)$ are
\begin{subequations}\small
\begin{align}
\deg_R A_n\Rlam &= \sum_{d|n} \varphi(n/d) 2^{d-1}, \\
\deg_R P_n\Rlam &\equiv \beta(n) = \sum_{d|n} \mu(n/d) 2^{d-1}, \\
\deg S_n(R) &= \beta(n) - \sum_{d | n, d < n} \beta(d) \, \varphi(n/d).
\end{align}
\label{eq:degR}
\end{subequations}
%\label{thm:degR}
%\end{theorem}
%
%
%
\begin{proof}
We first prove \refeqsub{degR}{a}.
%
We recall the subscript $p$ of $C_p$ denotes a sequence of indices $k$
  in the generating monomial $\prod_k {x_k}^{e_k}$.
But for a square-free cyclic polynomial,
  each $k$ occurs no more than once,
so $p$ also represents a \emph{set} of indices,
%
e.g.,
$p = 1$ represents $\{1\}$,
  ($C_1 = x_1 + \dots + x_n$, generator: $x_1$)
and
$p = 13$ represents $\{1, 3\}$
  ($C_{13} = x_1 \, x_3 + x_2 \, x_4 + \dots + x_n \, x_2$,
  generator: $x_1 \, x_3 $, assuming $n \ge 4$).
%more examples are listed in \reftab{sqrfreepoly}).
%
In this proof,
we shall also use $p$ to denote the corresponding index set,
  $|p|$ the set size, i.e., the number of indices in the set,
  and
  $\bar p \equiv \{1,\ldots,n\} \backslash p$ the complementary set.
%
Obviously, $|\bar p| + |p| = n$.
%
Further, we will include $p$ that correspond to alternative generators
  of the same cyclic polynomial,
e.g., we allow $p = \{2\}$, $\{3\}$, \dots, $\{n\}$,
  although they represent the same cyclic polynomial $C_1$
  as $p = \{1\}$.

Next, we recall the matrix elements $T_{pq}(R)$ arise from the square-free reduction
of $\Lambda(\vx) C_p(\vx) = (-2)^n x_1 \dots x_n \, C_p(\vx)$.
%
A single replacement ${x_k}^2 \rightarrow R - x_{k+1}$
produces two new terms:
in the first, ${x_k}^2 \rightarrow R$,
and in the second, ${x_k}^2 \rightarrow - x_{k+1}$.
We call the two type 1 and type 2 replacements, respectively.
%
If a monomial term $t(R, \vx)$
results from $l_1$ type 1 and $l_2$ type 2 replacements
during the reduction of a term $s(\vx)$ in $\Lambda(\vx) C_p(\vx)$,
then the degrees in $\vct x$, for any $x_k$, of $s(\vx)$ and $t(R, \vx)$
are related as
%
\begin{equation}
  \deg_\vx s(\vx) - \deg_\vx t(R, \vx) = 2 l_1 + l_2.
\label{eq:degxdiffst}
\end{equation}
%
Similarly, the degrees in $R$ satisfy
\[
  \deg_R s(\vx) - \deg_R t(R, \vx) = - l_1,
\]
but since $\deg_R s(\vx) = 0$,
\begin{equation}
  \deg_R t(R, \vx) = l_1.
\label{eq:degRt}
\end{equation}

%
Now if the monomial $t(R, \vx)$ settles in the $q$th column of
  the matrix $\vct T(R)$, as part
  of $T_{pq}(R) C_q(\vx)$ in \refeq{xcp0},
then
  $t(R, \vx)$
  must be
%the product of
%  a monomial of $R$ \big[as part of $T_{pq}(R)$\big]
%  and one of the
  a generator of $C_q(\vx)$;
so
%
\begin{equation}
  \deg_\vx t(R, \vx) = |q|.
\label{eq:degxt}
\end{equation}
%
Since $s(\vx)$ is part of $\Lambda(\vx) C_p(\vx)$,
  we have
%
\begin{equation}
  \deg_\vx s(\vx) = \deg_\vx \Lambda(\vx) + \deg_\vx C_p(\vx)
  = n + |p|.
\label{eq:degxs}
\end{equation}
%
%
From \refeqs{degxdiffst}, \req{degxt}, \req{degxs}, we get
\[
  n + |p| - |q| = 2 \, l_1 + l_2,
\]
and
\begin{equation}
  l_1  =    (n + |p| - |q| - l_2)/2
             \le  (n + |p| - |q|)/2.
\label{eq:l1limit}
\end{equation}
By \refeq{degRt}, we get
\begin{align*}
  \deg_R T_{pq}(R)
  = \max\{ \deg_R t(R, \vx) \}
  = \max\{ l_1 \}
  \le   (n + |p| - |q|)/2,
\end{align*}
where the equality holds when all replacements are type 1 ($l_2 = 0$).
%
Finally, each term of the determinant
$A_n\Rlam = \big|\lambda \, \vct I - \vct T(R)\big|$
is given by
$(-1)^s \prod_{p} \big[\lambda \, \delta_{p q} - T_{pq}(R)\big]$,
where $p$ runs through rows of the matrix
  and $\{q\}$ is a permutation of $\{p\}$,
  with $(-1)^s$ being the proper sign.
%
Summing over rows under this condition yields
\begin{align*}
  \deg_R A_n\Rlam
  = \max \sum_{p \in \Pset} \deg T_{pq}(R)
  \le \NB \, \big(n + |p| - |q|\big)/2 = n \NB/2,
\end{align*}
where equality can be achieved if $q = \bar p$ in every row.
By \refeq{necklace} we have \refeqsub{degR}{a}.
%
\refeqsub{degR}{b} and (c) follow from taking taking the degree in $R$ of
  \refeq{prod} and \refeq{origfac}.
\refeqsub{degR}{b} and (\ref{eq:degR}c) were known \cite{mira, stephenson1, blackhurst}.
%
\end{proof}
%
The first few $\deg_R A_n\Rlam$,
  $\deg_R P_n\Rlam$,
  and $\deg S_n(R)$ are
  1, 3, 6, 12, 20, 42, 70, 144, 270, 540, \dots,  %1034, 2112, 4108, \dots,
  1, 1, 3, 6, 15, 27, 63, 120, 252, 495, \dots, %1023, 2010, 4095, \dots,
  1, 0, 1, 3, 11, 20, 57, 108, 240, 472, \dots, % 1013, 1959, 4083, \dots,
  respectively, from $n = 1$.

@@

Let us summarize the algorithm.
%
First, we list \refeqs{xcp}  with $\Lambda(\vx) = \prod_{k=1}^n f'(x_k)$.
This step populates elements of the matrix $\vct T$.
%
The characteristic equation $A_n(\lambda) = \big|\lambda \, \vct I - \vct T\big| = 0$,
with $\lambda$ being $+1$ and $-1$, then gives
the polynomials at onset and bifurcation points, respectively.
%
To filter out the contributions from shorter $d$-cycles with $d|n$,
  we repeat compute the characteristic equations for all divisors $d$
  and then apply \req{primfac}.
%

When implemented on a computer,
it is often helpful to evaluate $A_n(\lambda)$ by Lagrange interpolation,
that is, we evaluate $A_n(r, \lambda)$ at a few different $r$,
e.g., $r = 0, \pm1, \pm2,\ldots$, then piece them together
to a polynomial.
The strategy also allows a trivial parallelization.
%
The algorithm was coded into a Mathematica program,
which was used to compute the polynomials for $n$ up to 13.
%
For polynomials of larger $n$ are saved in in the website in \refsec{end}.
%
The representative $r$ values are listed in \reftab{rval}.
%

The algorithm (implemented as a Mathematica program)
was quite efficient.
%
For the logistic map, the bifurcation point for $n = 8$
took three seconds to compute on a desktop computer
(single core, Intel\textsuperscript{\textregistered} Dual-Core CPU 2.50GHz).
%
In comparison,
  the same problem took roughly 5.5 hours \cite{kk1}
  using Gr\"obner basis
  and 44 minutes in a later study \cite{lewis}.
To be fair, using the latest Magma, computing the Gr\"obner basis
  took 81 and 14 minutes, on the same machine
  for \refeq{logmap} and \refeq{logmaps}, respectively;
even so, our approach still had a 200-fold speed-up.

%
The exact polynomials of these maps are generally too large to print on paper,
  e.g., the polynomial for the logistic map with $n = 13$
    takes roughly seven megabytes to write down.
%
We therefore save the polynomials and programs of the three maps on the web site
http://logperiod.appspot.com.

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
we recover the logistic map \refeq{logmaps}.

@@

As an example of a higher-order %one-dimensional
map, we consider the following cubic map \cite{strogatz}
%
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
This map is closely related to the one studied by Brown \cite{brown3, brown4}.
%
Unlike the logistic map, this map is antisymmetric,
  that is, $f(-x) = -f(x)$.
We shall show below that this odd parity can be used to reduce the amount of calculation.

@@

The characteristic polynomial $A_n(r, \lambda)$
  from the determinant equation
  has a degree $N_e(n)$ in $\lambda$.
Again, it encompasses the factors
  from the shorter $d$-cycles, as long as $d|n$.
The polynomial for the $n$-cycles can be obtained
  by \refeq{primfac} with proper substitutions, e.g.,
  $R\rightarrow r$.
%The degree of the polynomial,
The degree in $\lambda$ of
the minimal polynomial $P_n(r, \lambda)$ of the $n$-cycles,
  after some algebra,
  is given by
%
%
%\begin{theorem}
%The degree in $\lambda$ of
%the minimal polynomial $P_n(r, \lambda)$ of the $n$-cycles is
\begin{equation}
  L_e(n) = \frac{1}{n} \sum_{cd = n} \mu(c)
    \left[
      1 + \odd(c)\frac{3^d-1}{2}
    \right],
\label{eq:cublyndon}
\end{equation}
%where $\odd(n)$ %$= [1 - (-1)^n]/2$
%is $1$ for an odd $n$,
%but $0$ for an even $n$.
%\label{thm:cublyndon}
%\end{theorem}
%
%
For the first few values,
  $L_e(n)$ = 2, 2, 4, 10, 24, 60, 156, 410, \ldots, starting from $n = 1$.
%
This is also the number of the $n$-cycles \cite{hao, hao2}.
Note that, for $n>1$, $x_k$ in half of the cycles
  with negative $r$ are imaginary.
However, in a closely-related map
$z_{k+1} = r \, z_k \, (1 - {z_k}^2)$,
which differs from \refeq{cubic} by $x_k = \sqrt{r} z_k$,
$z_k$ in the negative-$r$ cycles are real.
%
%
Following similar steps in \refsec{degR},
we find the corresponding degrees in $r$
of the characteristic polynomial $A_n(r, \lambda)$
  and minimal polynomial $P_n(r, \lambda)$ of the $n$-cycles
are $n N_e(n)$ and $n L_e(n)$, respectively.

@@

The symmetry $f(-x) = -f(x)$ permits
  a particular \emph{odd-cycle},
  see \reffig{oddcycle} for examples.
If $x_{n+1} = -x_1$ has a solution,
  then $x_1, \ldots, x_{n}, -x_1, \ldots, -x_{n}$
  is a $2n$-cycle, for $x_{2n+1} = - x_{n+1} = x_1$.
%
The existence of odd-cycles is why we used even, but not odd, cyclic polynomials
  in the solution.
An odd $2n$-cyclic polynomials is also always zero
  in an $n$-odd-cycle,
e.g., $C_1 = x_1 + x_2 + \dots + x_n - x_1 - x_2 - \dots - x_n = 0$;
and the zero determinant condition \refeq{xr}
  would omit the $n$-odd-cycles from the solution.

@@

\nonumsection{Acknowledgments} \noindent
I thank Drs. T. Gilbert and Y. Mei for helpful communications.
Computing time on the Shared University Grid at Rice,
funded by NSF under Grant EIA-0216467, is gratefully acknowledged.

>>>>

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx_{#1}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\lam}{(\lambda)}
\newcommand{\Rlam}{(R, \lambda)}
\newcommand{\rlam}{(r, \lambda)}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = f(x_k) = x_{k+1}$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  the starting point of the next iteration.
%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle},
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if a small deviation from $x^*$
  gets reduced by successive iterations.
%
For a differentiable $f$,
  this is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz}.

@@

The $n$-cycle polynomial
  $P_n(R, \lambda)$
  can be further factorized at $\lambda = 1$.
%
This is caused by the intersection of an $n$-cycle
  and a shorter $d$-cycle ($d|n$, $d < n$),
  which forces the two to share orbit.
%
One can show that (1) such an intersection happens only
  at the onset of the longer $n$-cycle, where $\lambda = 1$,
%
  and (2) $d$ and $n$ are coprime \cite{blackhurst}.
%
The simplest example is the first bifurcation point
  at $R = 3/4$ for $d = 1, n = 2$,
  where the fixed point \refeq{xr1}
  undergoes period-doubling and becomes the 2-cycle \refeq{xr2}.
As a result, the $n$-cycle polynomial $P_n(R, \lambda)$
  has to accommodate the $d$-cycle one $P_d(R, \lambda')$ as a factor,
  with $\lambda'$ being a primitive $(n/d)\,$th root of $\lambda = 1$.

We can therefore distinguish \emph{original} cycles
  from those born out of the above branching.
Similar to \refeq{QP},
  the contributions from the non-original cycles
  born out of the $d$-cycles
  can be written as a cyclotomic product
  $\prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)$,
where $k$ runs through integers from 1 to $c=n/d$,
  that are coprime to $c$.
With these contributions removed,
  we obtain the polynomial $S_n(R)$ responsible for the original cycles:
  %or the \emph{original factor} below,
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      %\left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      %\right]
    },
  \label{eq:origfac}
  \end{equation}
%
%
%
Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$.
%  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
%  the onset of the original cycle.
Note $R_a' = 5/4$ is not original,
  for it is born out of period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).

@@

In the numerical implementation
  (see \refsec{end}),
  it is helpful to know the degrees in $R$ of the polynomials
  to evaluate the determinant by Lagrange interpolation. 
We first show that:
\begin{equation}
\deg_R A_n\Rlam = \textstyle\sum_{d|n} \varphi(n/d) 2^{d-1},
\label{eq:degRA}
\end{equation}
%
The first few values are
1, 3, 6, 12, 20, 42, 70, 144, 270, 540, \dots, from $n =1$.%1034, 2112, 4108, \dots,

From \refeq{xr},
  every term in the expansion of the determinant $|\vct A|$
  can be written as
  $\prod_{p} A_{p,q}$,
  in which, $p$ runs through all cyclic polynomials,
  and $\{q = q(p)\}$ is a permutation of $\{p\}$.
Thus, $\deg_R |\vct A|$ is
  reached by the term that maximizes the product.
Recall that $A_{p,q}$ results from reducing
  $\Lambda(\vx) \, C_p(\vx)$ to $C_q(\vx)$
  by a sequence of substitutions ${x_k}^2 = R-x_{k+1}$.
It follows that to increase $\deg_R$ by 1,
  $\deg_\vx$ is decreased by at least 2.
Accordingly,
  $\deg_R A_{p,q} \le \deg_\vx(\Lambda \, C_p) - \deg_\vx C_q$.
Now running $p$ and $q$ through the $\NB$ cyclic polynomials
  yields
\[
  \deg_R |\vct A | = \deg_R \textstyle\prod_{p} A_{p,q} \le n \, \NB(n),
\]
where we have used $\deg_\vx \Lambda(\vx) = n$.
%
The equality can be achieved, if each $q$ is the ``complement'' of $p$,
  that is, if $\prod_k x_k^{e_k}$ is the generator of $C_p$,
  $C_q$ can be generated from $\prod_k x_k^{1-e_k}$.
For example, for $n=4$,
  $C_1 = x_1 + \cdots + x_4$
  is the complement of
  $C_{123} = x_{234} + \cdots + x_{123}$;
  while $C_{12}$ is its own complement.
Finally, we reach \refeq{degRA} by \refeq{necklace}.

Now, taking the degrees in $R$ of \refeq{primfac} and \refeq{origfac}
  yields
%
\begin{subequations}
\begin{align}
\deg_R P_n\Rlam &\equiv \beta(n) = \sum_{d|n} \mu\Big(\frac{n}{d}\Big) 2^{d-1}, \\
    \deg S_n(R) &= \beta(n)
  - \sum_{d | n, d < n} \beta(d) \, \varphi\Big(\frac{n}{d}\Big).
\end{align}
\label{eq:degR}
\end{subequations}
%
The two formulas are known \cite{mira, stephenson1, blackhurst}.
%
%
The first few
  $\deg_R P_n\Rlam$,
  and $\deg S_n(R)$ are
  1, 1, 3, 6, 15, 27, 63, 120, 252, 495, \dots, %1023, 2010, 4095, \dots,
  1, 0, 1, 3, 11, 20, 57, 108, 240, 472, \dots, % 1013, 1959, 4083, \dots,
  respectively, from $n = 1$.

@@

Let us summarize the algorithm.
%
First, we list \refeqs{xcp}  with $\Lambda(\vx) = \prod_{k=1}^n f'(x_k)$.
This step populates elements of the matrix $\vct T$.
%
The characteristic equation $A_n(\lambda) = \big|\lambda \, \vct I - \vct T\big| = 0$,
with $\lambda$ being $+1$ and $-1$, then gives
the polynomials at onset and bifurcation points, respectively.
%
The contributions from shorter $d$-cycles
  can be removed by
  % computing the characteristic equations for all divisors $d$ and applying
  \refeq{primfac}.
%

When implemented on a computer,
it is often helpful to evaluate the characteristic polynomial
by Lagrange interpolation,
that is, by evaluating $A_n\Rlam$ at a few different $R$,
e.g., $R = 0, \pm1, \pm2,\ldots$, then piecing them together
to a polynomial.
The strategy also allows a trivial parallelization.
%
The algorithm was coded into a Mathematica program,
which was used to compute the polynomials for $n$ up to 13.
%
As shown in \reffig{logtime}, the algorithm
outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods,
  by at least two orders of magnitudes.
%
Note that the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the matrix of the characteristic equation \refeq{xr}
  is of only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
Thus, the current algorithm is expected to be faster.

\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{logtime.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:logtime}
  Time to compute the $n$-cycle bifurcation polynomials,
    measured from a single core of
    Intel\textsuperscript{\textregistered} 2.5GHz Dual-Core CPU
  (Mathematica version: 8, Magma version: 2.18-4).
  }
\end{figure}
%

%
The exact polynomials of these maps are generally too large to print on paper,
  e.g., the polynomial for the logistic map with $n = 13$
    takes roughly seven megabytes to write down.
%
We therefore save the polynomials and programs on the web site
http://logperiod.appspot.com.
%
The representative $r$ values are listed in \reftab{rval}.

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
%\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
we recover the logistic map \refeq{logmaps}.

@@

As an example of a higher-order %one-dimensional
map, we consider the following cubic map \cite{strogatz}
%
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
%
Unlike the logistic map, this map is antisymmetric,
  that is, $f(-x) = -f(x)$.
%
In fact, all antisymmetric cubic maps \cite{brown3, brown4}
  can be mapped to \refeq{cubic} by a change of variables.
We shall show below that this odd parity can be used to reduce the amount of calculation.

@@

The characteristic polynomial $A_n\rlam$
  from the determinant equation
  has a degree $N_e(n)$ in $\lambda$.
Again, it encompasses the factors
  from the shorter $d$-cycles ($d|n$),
  and the $n$-cycle polynomial can be obtained
  from \refeq{primfac} by proper substitutions, e.g.,
  $R\rightarrow r$.
%The degree of the polynomial,
The degree in $\lambda$ of $P_n\rlam$
  is given by, after some algebra,
%
%
\begin{equation}
  L_e(n) = \frac{1}{n} \sum_{cd = n} \mu(c)
    \left[
      1 + \odd(c)\frac{3^d-1}{2}
    \right],
\label{eq:cublyndon}
\end{equation}
%
%
For the first few values,
  $L_e(n)$ = 2, 2, 4, 10, 24, 60, 156, 410, \ldots, starting from $n = 1$.
%
This is also the number of the $n$-cycles \cite{hao, hao2}.
Note that, for $n>1$, $x_k$ in half of the cycles
  with negative $r$ are imaginary.
However, in a transformed map
$z_{k+1} = r \, z_k \, (1 - {z_k}^2)$,
which differs from \refeq{cubic} by $x_k = \sqrt{r} z_k$,
$z_k$ in the negative-$r$ cycles are real.
%
%
Following similar steps in \refsec{degR},
we find that
  $\deg_r A_n(r, \lambda) = n N_e(n)$
  and $\deg_r P_n(r, \lambda) = n L_e(n)$.

@@

The symmetry $f(-x) = -f(x)$ permits
  a particular \emph{odd-cycle},
  see \reffig{oddcycle} for examples.
If $x_{n+1} = -x_1$ has a solution,
  then $x_1, \ldots, x_{n}, -x_1, \ldots, -x_{n}$
  is a $2n$-cycle, for $x_{2n+1} = - x_{n+1} = x_1$.
%
It follows that a cyclic polynomial
  formed by the $2n$ points
  with an odd degree in $\vxn{2n}$
  is also always zero,
  e.g., $C_1 = x_1 + x_2 + \dots + x_n - x_1 - x_2 - \dots - x_n = 0$.
Accordingly, the characteristic equation \refeq{xr}
  would omit the $n$-odd-cycles from the solution.
The existence of odd-cycles is why we use even, but not odd,
  cyclic polynomials in the solution.

@@

Finally, we mention that the $n$-cycle polynomial $P_n\rlam$
  is an even polynomial of $r$ if $n$ is a multiple of 4.
%This follows from that the transformation
%  $x_k \equiv (-)^k i y_k, r \equiv -r'$
%  yields the identical map: $y_{k+1} = r' y_k - {y_k}^3$.
Practically, we can cut the computating time by half
  when calculating the determinant by Lagrange interpolation
  in this case.

\nonumsection{Acknowledgments} \noindent
It is a pleasure to thank T. Gilbert, Y. Mei and the referees
  for helpful communications and suggestions.
Computing time on the Shared University Grid at Rice,
  funded by NSF under Grant EIA-0216467, is gratefully acknowledged.
